---
title: "Data Visualization"
author: "Luiz and Morgan"
date: "2024-06-08"
output: html_document
---

# Loading libraries
```{r}
library(here)
library(lme4)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(multcomp)
library(broom)
library(viridis)
library(Matrix)
# for later plotting
library(cowplot)
library(scales)


```





# DMS Data preprocessing
```{r}

data <- read.csv(here('results/compression/summary_DMS_esm2_150M_all_methods.csv'))
data <- dplyr::select(data, Dataset, Compression_method, Fold, R2_score_test, rho_score_test)


data

# Convert categorical variables to factors and releveling the mean to be the reference group
data$Compression_method <- factor(data$Compression_method)
data$Compression_method <- relevel(data$Compression_method, ref = "mean")
data$Dataset <- factor(data$Dataset)
data$Fold <- factor(data$Fold)
data
```

```{r}

#data <- data %>%
#  mutate(Dataset = case_when(
#     dataset == "HG_FLU_Bloom2016" ~ "New_A",
#     dataset == "BLAT_ECOLX_Ostermeier2014" ~ "New_B",
#     dataset == "RL401_YEAST_Bolon2014" ~ "New_C",
#     dataset == "PABP_YEAST_Fields2013_doubles" ~ "New_C",
#     dataset == "B3VI55_LIPST_Whitehead2015" ~ "New_C",
#     dataset == "TIM_THETH" ~ "New_C",
#     dataset == "CALM1_HUMAN_Roth2017" ~ "New_C",
#     dataset == "parEparD_Laub2015_all" ~ "New_C",
#     dataset == "BF520_env_Bloom2018" ~ "New_C",
#     dataset == "a" ~ "New_C",
#     dataset == "a" ~ "New_C",
#     dataset == "a" ~ "New_C",
#     
#     old_names = ['',
#        '', '',
#        '', '', '', 'BF520_env_Bloom2018',
#        'UBC9_HUMAN_Roth2017', 'DLG4_RAT_Ranganathan2012',
#        'BRCA1_HUMAN_RING', 'TPMT_HUMAN_Fowler2018',
#        'MTH3_HAEAESTABILIZED_Tawfik2015', 'HSP82_YEAST_Bolon2016',
#        'MK01_HUMAN_Johannessen', 'PTEN_HUMAN_Fowler2018',
#        'BG505_env_Bloom2018', 'BLAT_ECOLX_Tenaillon2013', 'TIM_SULSO',
#        'B3VI55_LIPSTSTABLE', 'RL401_YEAST_Bolon2013',
#        'PABP_YEAST_Fields2013_singles', 'BLAT_ECOLX_Ranganathan2015',
#        'RL401_YEAST_Fraser2016', 'TIM_THEMA', 'AMIE_PSEAE_Whitehead',
#        'BG_STRSQ_hmmerbit', 'GAL4_YEAST_Shendure2015',
#        'BLAT_ECOLX_Palzkill2012', 'SUMO1_HUMAN_Roth2017',
#        'RASH_HUMAN_Kuriyan', 'YAP1_HUMAN_Fields2012_singles',
#        'KKA2_KLEPN_Mikkelsen2014', 'TPK1_HUMAN_Roth2017',
#        'PA_FLU_Sun2015', 'IF1_ECOLI']
# new_names = ['HG FLU', 'BLAT ECOLX 2014', 'RL401 2014',
#        'PABP doubles', 'B3VI55 LIPST',
#        'TIM THETH', 'CALM1 HUMAN', 'parEparD all',
#        'BF520 env', 'UBC9 HUMAN', 'DLG4 RAT',
#        'BRCA1 RING', 'TPMT HUMAN',
#        'MTH3 HAEAESTABILIZED', 'HSP82 YEAST',
#        'MK01 HUMAN', 'PTEN HUMAN',
#        'BG505 env', 'BLAT ECOLX 2013', 'TIM SULSO',
#        'B3VI55 LIPSTSTABLE', 'RL401 2013',
#        'PABP singles', 'BLAT ECOLX 2015',
#        'RL401 2016', 'TIM THEMA', 'AMIE PSEAE',
#        'BG STRSQ', 'GAL4 YEAST',
#        'BLAT ECOLX 2012', 'SUMO1 HUMAN',
#        'RASH HUMAN', 'YAP1 singles',
#        'KKA2 KLEPN', 'TPK1 HUMAN',
#        'PA FLU', 'IF1 ECOLI']
#     
#     TRUE ~ Dataset  # Default case if no match
#   ))

```


# DMS Fitting the lme4 model

Dependent Variable: The dependent variable (response). Here is R2_score_test.       

Fixed Effect: The predictor variables (fixed effects) can be numeric or categorical. Categorical variables should be factors.   
'Compression_methd' is included as a fixed effect to see how different compression methods influence the R2_score_test.     

Random Effects: The grouping variables (random effects) should be factors. In here are Dataset and Fold.   
Both Dataset and Fold are included as random effects to account for variability across datasets and cross-validation folds.  

```{r}
model <- lmer(R2_score_test ~ Compression_method + (1 | Dataset) + (1|Fold), data=data)
#model <- lmer(R2_score_test ~ Compression_methd + (1 | Dataset), data=data)
########model <- lmer(rho_score_test ~ Compression_methd + (1 | Dataset) + (1|Fold), data=data)
#model <- lmer(rho_score_test ~ Compression_methd + (1 | Dataset), data=data)
summary(model)



# if you get this error:
# Error in initializePtr() : 
#  function 'chm_factor_ldetL2' not provided by package 'Matrix'

#visit: https://www.bioconductor.org/packages/release/bioc/vignettes/dreamlet/inst/doc/errors.html
# look at 'Errors with random effects' section
```



```{r}
#glht_results <- summary(glht(model, linfct=c('Compression_methdbos=0',
#                             'Compression_methdiDCT=0',
#                              'Compression_methdmaxPool=0',
#                              'Compression_methdpca1=0',
#                              '`Compression_methdpca1-2`=0',
#                              'Compression_methdpca2=0',
#                              'Compression_methdrbf1=0',
#                              'Compression_methdrbf2=0',
#                              'Compression_methdsigmoid1=0',
#                              'Compression_methdsigmoid2=0')))



glht_results <- summary(glht(model, linfct=c(
                              'Compression_methodbos=0',
                              'Compression_methodiDCT1=0',
                              'Compression_methodiDCT2=0',
                              'Compression_methodiDCT3=0',
                              'Compression_methodiDCT4=0',
                              'Compression_methodiDCT5=0',
                              'Compression_methodmaxPool=0',
                              'Compression_methodpca1=0',
###                              '`Compression_methodpca1-2`=0',
                              'Compression_methodpca2=0',
                              'Compression_methodrbf1=0',
                              'Compression_methodrbf2=0',
                              'Compression_methodsigmoid1=0',
                              'Compression_methodsigmoid2=0')))

glht_results 
```


```{r}
#summary(glht(model, linfct=mcp(Compression_method="Tukey")))
```


## Interpretation


The intercept in a regression model represents the expected value of the dependent variable (in this case, R2_score_test) 
when all predictor variables are set to zero. It's the baseline value from which other effects are measured.


Fixed effects are the estimated coefficients for the predictor variables in the model. These effects are assumed to be the 
same across all groups in the dataset. In your model, Compression_methd is treated as a fixed effect, meaning its impact on 
R2_score_test is consistent across all datasets.

Random effects account for variability within the data that is not explained by the fixed effects. They allow the model to 
consider differences across groups (e.g., different datasets or cross-validation folds). For example, if some datasets inherently
have higher or lower R2 scores, this variability is captured by including Dataset as a random effect.

Residuals are the differences between the observed values and the values predicted by the model. They represent the "error" 
or "noise" in the model. Residual analysis helps to check the assumptions of the model (e.g., normality and homoscedasticity 
of residuals).



# DMS Evaluating individual datasets

```{r}
# Tidy the results
tidy_glht <- tidy(glht_results)
#tidy_glht<- tidy_glht %>% dplyr::filter(contrast != #'Compression_methodiDCT')

ci <- tidy(confint(glht_results)) %>% dplyr::select("contrast", "conf.high", "conf.low")

res_glht = merge(tidy_glht, ci, by = 'contrast')



res_glht$contrast <- gsub("Compression_method", "", res_glht$contrast)

#res_glht$contrast <- gsub("iDCT2", "iDCT", res_glht$contrast)


# Determine significance levels
res_glht$signif <- ifelse(res_glht$adj.p.value < 0.001, "***",
                 ifelse(res_glht$adj.p.value < 0.01, "**",
                 ifelse(res_glht$adj.p.value < 0.05, "*", "")))



res_glht
```



# Figure 2A: DMS Plotting lmer
```{r}
#, fig.width=2.5, fig.height=6, out.width = 100%, out.height = 100%


res_glht <- res_glht %>%
  mutate(contrast = case_when(
    contrast == "bos" ~ "BOS",
    contrast == "maxPool" ~ "Max Pooling",
    contrast == "rbf1" ~ "RBF PC1",
    contrast == "rbf2" ~ "RBF PC2",
    contrast == "sigmoid1" ~ "Sigmoid PC1",
    contrast == "sigmoid2" ~ "Sigmoid PC2",
    contrast == "pca1" ~ "PC1",
    contrast == "pca2" ~ "PC2",
    
    TRUE ~ contrast
  ))

final_plot <- res_glht %>%
  ggplot(aes(y = fct_reorder(contrast, estimate), 
             x = estimate, 
             #color = contrast
             )) +
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2) +
  theme_minimal() +
  labs(
       y = "Compression Method",
       x = expression(R^2 * " Difference from Mean Pooling")
       ) +
  geom_vline(xintercept = 0, linetype = "dotted", color = '#000000', linewidth = 1) +

  scale_x_continuous(limit = c(-.25, 0.05),
  ) + 
  theme_minimal_vgrid(12, rel_small = 1) +  
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    plot.background = element_rect(fill = "white", colour = "white"),
    axis.text.x=element_text(size=12),
    axis.text.y=element_text(size=10),

    panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
    axis.line = element_blank(),
    axis.ticks.x = element_line(color = "black"),
    axis.ticks.y = element_line(color = "black"),
    plot.margin = margin(10,10,0,10)
  ) +
  ggtitle("DMS")

 #ggsave(final_plot, 
 #       filename = "fig2_iDCT_DMS_v12.png",
 #       path = here("figures"),
 #       width = 6,
 #       height = 6,
 #       dpi=600)


final_plot
  
```



# Fig 3A DMS multi boxplot 
```{r}
multi_mod_DMS <- read.csv(here("results/lassoCV/DMS/DMS_all_multi_model_metadata.csv"))

multi_mod_DMS <- multi_mod_DMS %>%
  mutate(Model = case_when(
    Model == "esm1v" ~ "ESM-1v",
    Model == "esm2" ~ "ESM-2",
    Model == "esmc" ~ "ESM C",
    Model == "amplify" ~ "AMPLIFY"
  )) %>%
  mutate(
    parameters = case_when(
      parameters == 6.5e8 & Model == "ESM-2" ~ 7.8e8,
      parameters == 6e8 & Model == "ESM C" ~ 5.5e8,
      TRUE ~ parameters
    )
  ) 

#exlude due to sequences too long for esm1v, need to make sup figure
exclude_data <- c("BRCA1 RING", "UBE4B singles", "BRCA1 BRCT", "POLG Sun2014")
#exclude_data <-  c('BRCA1_HUMAN_RING','UBE4B_MOUSE_Klevit2013_singles', 'BRCA1_HUMAN_BRCT', 'POLG_HCVJF_Sun2014')

grouped_multi_mod_DMS <- multi_mod_DMS %>%
  filter(!Dataset_Name %in% exclude_data) %>%
  filter(Compression_method %in% c('mean')) %>%
  group_by(Dataset_Name, Model, parameters) %>%
  summarise(
    mean_score = mean(get('R2_score_test'), na.rm = TRUE),
    std_score = sd(get('R2_score_test'), na.rm = TRUE),
   
  ) %>%
  mutate(Model = factor(Model, levels = c("ESM-1v", "ESM-2", "ESM C", "AMPLIFY"))) %>%
  ungroup()



data_medians <- grouped_multi_mod_DMS %>%
  group_by(Model, parameters) %>%
  summarize(
    median_R2 = median(mean_score)
  )



# need to make an order for the x axis
#grouped_multi_mod_DMS$Model <- factor(grouped_multi_mod_DMS$Model, levels = c("ESM1v 650M", "ESM2 8M", "ESM2 35M", "ESM2 150M", "ESM2 650M", "ESM2 3B", "ESM2 15B"))


```


```{r}


DMS_all_boxplot <- grouped_multi_mod_DMS %>%
  ggplot(aes(
    x = parameters, 
    y = mean_score, 
    group = interaction(Model, parameters), 
    color = Model,
    fill = Model
    )) +
    geom_line(
      data = data_medians,
      aes(parameters, median_R2, color = Model),
      inherit.aes = FALSE
    ) +
    geom_boxplot()+
    scale_x_log10(trans = scales::log_trans(),
                     breaks = scales::log_breaks(),
                     labels = trans_format("log10", math_format(10^.x))) +
    
    #scale_color_manual(values = color_dict) +
    scale_color_manual(
      values = colorspace::darken(c(`ESM-1v` = "#000000", `ESM-2` = "#4292C6",
                 `ESM C` = "#228C22", AMPLIFY = "#ff7f0e"), .2)
    ) +
    scale_fill_manual(
      values = c(`ESM-1v` = "#00000080", `ESM-2` = "#9ECAE1",
                 `ESM C` = "#5ad75a", AMPLIFY = "#ffb16c")
    ) +
    labs(x = "", 
         y = expression(paste("Cross Validation ", R^2)), 
         #title = ""
         color = ""
         ) +
  guides(
    color = guide_legend(title = "Model"),
    fill = guide_legend(title = "Model")
  ) +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10))
  
DMS_all_boxplot
 
# ggsave(DMS_all_boxplot,
#       filename = "results/lassoCV/DMS/fig3_multi_model.png",
#        path = here('figs'),
#        width = 8,
#        height = 4,
#       dpi=600)



```


# Sup 3 excluded
```{r}
exclude_data <- c("BRCA1 RING", "UBE4B singles", "BRCA1 BRCT", "POLG Sun2014")
#exclude_data <-  c('BRCA1_HUMAN_RING','UBE4B_MOUSE_Klevit2013_singles', 'BRCA1_HUMAN_BRCT', 'POLG_HCVJF_Sun2014')

exclude_multi_mod_DMS <- multi_mod_DMS %>%
  filter(Dataset_Name %in% exclude_data) %>%
  filter(Compression_method %in% c('mean')) %>%
  group_by(Dataset_Name, Model, parameters) %>%
  summarise(
    mean_score = mean(get('R2_score_test'), na.rm = TRUE),
    std_score = sd(get('R2_score_test'), na.rm = TRUE),
    prot_length = median(get('prot_length'))
   
  ) %>%
  mutate(Model = factor(Model, levels = c("ESM-1v", "ESM-2", "ESM C", "AMPLIFY"))) %>%
  ungroup()

max_len = max(exclude_multi_mod_DMS$prot_length)
max_len = min(exclude_multi_mod_DMS$prot_length)

```


```{r}


excluded_sup3 <- exclude_multi_mod_DMS %>%
  ggplot(aes(
    x = parameters, 
    y = mean_score, 
    group = Dataset_Name,
    color = Dataset_Name
    #fill = Model
    )) +
    geom_line() +
    geom_point(shape = 21, 
               size = 2,  
                 color = "black", 
                 aes(fill = Dataset_Name),
                 stroke = .2
                 ) + 
    
    geom_ribbon(aes(
      ymin = mean_score - std_score,
      ymax = mean_score + std_score, fill = Dataset_Name), alpha = 0.2, color = NA) +
  
      scale_x_log10(trans = scales::log_trans(),
                       breaks = scales::log_breaks(),
                       labels = trans_format("log10", math_format(10^.x))) +
      
      labs(x = "Parameters", 
           y = expression(paste("Cross Validation ", R^2)), 
           #title = "Dataset",
           #color = ""
           ) +
        guides(
        #color = guide_legend(override.aes = list(shape = 16, fill = NA)),
        color = guide_legend(title = "Dataset"),
        
        fill = guide_legend(title = "Dataset"))+
          #override.aes = list(shape = 16, color = NA))) +
    
      theme_minimal_grid(12, rel_small = 1)+
      theme(
          legend.position = "top",
          panel.background = element_rect(fill = "white", colour = "white"),
          plot.background = element_rect(fill = "white", colour = "white"),
        
  
          panel.border = element_rect(color = "black", 
                                      fill = NA, 
                                      size = .5),
          axis.line = element_blank(),
          axis.ticks.x = element_line(color = "black"),
          axis.ticks.y = element_line(color = "black"),
          plot.margin = margin(10, 10, 0, 10))+
      facet_wrap(~ Model, ncol = 1) 
    
excluded_sup3
 
# ggsave(excluded_sup3,
#       filename = "excluded_sup3.png",
#        path = here('figs'),
#        width = 7.5,
#        height = 8,
#       dpi=600)


```









# Figure 5, sup 6, 7, 8 


```{r}

data_with_size <- read.csv(here('results/lassoCV/DMS/DMS_all_multi_model_metadata.csv'))

data_with_size <- data_with_size %>%
  filter(Compression_method == "mean") %>%
  group_by(Dataset, Model, ModelSize) %>%
  summarise(
    mean_score = mean(get('R2_score_test'), na.rm = TRUE),
    std_score = sd(get('R2_score_test'), na.rm = TRUE),
    dataset_size = median(get('dataset_size')),
    group = unique(group)[1],
    prot_length = median(get('prot_length'))
  ) %>%
  mutate(Model = case_when(
    Model == "esm1v" ~ "ESM1v",
    Model == "esm2" ~ "ESM2",
    Model == "esmc" ~ "ESMC",
    Model == "amplify" ~ "Amplify"
  ))%>%
  mutate(Model = factor(Model, levels = c("ESM1v", "ESM2", "ESMC", "Amplify"))) %>%
  ungroup()


#data_with_size$Model <- factor(data_with_size$Model, levels = c('ESM1v 650M', 'ESM2 8M', 'ESM2 35M', 'ESM2 150M', 'ESM2 650M', 'ESM2 3B', 'ESM2 15B'))
data_with_size$group <- factor(data_with_size$group, levels = c("Growth", "Viral replication", "Enzyme function", "Other", "Peptide binding", "Protein stability"))

data_with_size$ModelSize <- factor(data_with_size$ModelSize, levels = c("8M", "35M", "120M", "150M", "350M", "650M","3B", "15B", "300M", "600M"))


```


# Sample Size
```{r}
#ESM1v 650M
model_ESM1v_650M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM1v" & ModelSize == "650M"))
r_squared_ESM1v_650M <- summary(model_ESM1v_650M)$r.squared
label_ESM1v_650M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_ESM1v_650M)))


# ESM2 8M
model_8M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM2" & ModelSize == "8M"))
r_squared_8M <- summary(model_8M)$r.squared
label_8M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_8M)))


#ESM2 35M
model_35M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM2" & ModelSize == "35M"))
r_squared_35M <- summary(model_35M)$r.squared
label_35M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_35M)))

#ESM2 150M
model_150M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM2" & ModelSize == "150M"))
r_squared_150M <- summary(model_150M)$r.squared
label_150M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_150M)))


#ESM2 650M
model_650M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM2" & ModelSize == "650M"))
r_squared_650M <- summary(model_650M)$r.squared
label_650M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_650M)))

#ESM2 3B
model_3B <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM2" & ModelSize == "3B"))
r_squared_3B <- summary(model_3B)$r.squared
label_3B <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_3B)))


#ESM2 15B
model_15B <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESM2" & ModelSize == "15B"))
r_squared_15B <- summary(model_15B)$r.squared
label_15B <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_15B)))


#ESMC 300M
model_C300M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESMC" & ModelSize == "300M"))
r_squared_C300M <- summary(model_C300M)$r.squared
label_C300M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_C300M)))


#ESMC 600M
model_C600M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESMC" & ModelSize == "600M"))
r_squared_C600M <- summary(model_C600M)$r.squared
label_C600M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_C600M)))

# #ESMC 6B
# model_C6B <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "ESMC" & ModelSize == "6B"))
# r_squared_C6B <- summary(model_C6B)$r.squared
# label_C6B <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_C6B)))


#Amplify 120M 
model_A120M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "Amplify" & ModelSize == "120M"))
r_squared_A120M <- summary(model_A120M)$r.squared
label_A120M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_A120M)))

#Amplify 350M
model_A350M <- lm(mean_score ~ log(dataset_size), data = subset(data_with_size, Model == "Amplify" & ModelSize == "350M"))
r_squared_A350M <- summary(model_A350M)$r.squared
label_A350M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", r_squared_A350M)))




data_text <- data.frame(
  labels = I(list(label_ESM1v_650M, label_8M, label_35M, label_150M, label_650M, label_3B, label_15B, label_C300M, label_C600M, label_A120M, label_A350M)),
  Model = c("ESM1v", "ESM2", "ESM2", "ESM2", "ESM2","ESM2", "ESM2", "ESMC", "ESMC", "Amplify", "Amplify"),
  ModelSize = c("650M", "8M", "35M", "150M", "650M","3B", "15B", "300M", "600M", "120M", "350M"),
  x = c(1150, 3700, 3700, 3700, 3700, 3700, 3700, 1200, 1200, 1150, 1150),
  y = c(.75, .75, .75, .75, .75, .75, .75, .83, .83, .76, .76)
)



data_text$ModelSize <- factor(data_text$ModelSize, levels = c("8M", "35M", "120M", "150M", "350M", "650M","3B", "15B", "300M", "600M"))



print(data_text)
```



```{r}

# FOR SUP 8 - ESM2

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, parse(text = paste0("10^", log10(x))), "")
}


plot_param <- data_with_size %>%
  filter(Model == 'ESM2') %>%
    ggplot(aes(
      x = dataset_size, 
      y = mean_score,
      color = group
      )) +

  scale_x_log10(labels = label_function)+



  geom_smooth(
      aes(group = Model),
      method = 'lm', 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
    geom_point(alpha = 1, size = .9
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .7
  
      )+
  
 
    labs(x = "Sample Size", 
         y = expression(paste("Cross Validation ", R^2))
         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        #legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10)
        )+
  
    geom_text(
        data = filter(data_text, Model == "ESM2"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,

        )+
  facet_wrap(~ ModelSize,scales = "free_x")
    


 # ggsave(plot_param,
 #        filename = "plot_param.png",
 #        path = here("figs"),
 #        width = 8.5,
 #        height = 5,
 #        dpi=600)

plot_param 

```


```{r}

#Sup 9 ESMC

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, parse(text = paste0("10^", log10(x))), "")
}

plot_param_ESMC <- data_with_size %>%
  filter(Model == 'ESMC' ) %>%
    ggplot(aes(
      x = dataset_size, 
      y = mean_score,
      color = group
      )) +
  geom_smooth(

      method = 'lm', 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
  
  scale_x_log10(labels = label_function)+

    scale_y_continuous(limits = c(0, 0.85))+
  
    geom_point(alpha = 1, size = 2
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .8,
      size = .6
      )+
    labs(x = "Sample Size", 
         y = expression(paste("Cross Validation ", R^2))
           #expression(paste(R^2)),

         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        legend.position = "none",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10)
        )+
  
    geom_text(
        data = filter(data_text, Model == "ESMC" & ModelSize == '300M'),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,
        
        )

plot_param_ESMC

# 
# ggsave(plot_param_ESMC,
#        filename = "plot_param_ESMC.png",
#        path = here("figs"),
#        width = 8,
#        height = 5,
#        dpi=600)

```

```{r}
# plot_param_Amplify

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, parse(text = paste0("10^", log10(x))), "")
}


plot_param_Amplify <- data_with_size %>%
  filter(Model == 'Amplify') %>%
    ggplot(aes(
      x = dataset_size, 
      y = mean_score,
      color = group
      )) +

  scale_x_log10(labels = label_function)+



  geom_smooth(
      aes(group = Model),
      method = 'lm', 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
    geom_point(alpha = 1, size = .9
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .7
  
      )+
  
 
    labs(x = "Sample Size", 
         y = expression(paste("Cross Validation ", R^2))
         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        #legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10)
        )+
  
    geom_text(
        data = filter(data_text, Model == "Amplify"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,

        )+
  facet_wrap(~ ModelSize,scales = "free_x")
    


 # ggsave(plot_param_Amplify,
 #        filename = "plot_param_Amplify.png",
 #        path = here("figs"),
 #        width = 8.5,
 #        height = 5,
 #        dpi=600)

plot_param_Amplify 



```




# Protein Length

```{r}

#ESM1v 650M
pmodel_ESM1v_650M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM1v" & ModelSize == "650M"))
pr_squared_ESM1v_650M <- summary(pmodel_ESM1v_650M)$r.squared
plabel_ESM1v_650M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_ESM1v_650M)))

# ESM2 8M
pmodel_8M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM2" & ModelSize == "8M"))
pr_squared_8M <- summary(pmodel_8M)$r.squared
plabel_8M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_8M)))

#ESM2 35M
pmodel_35M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM2" & ModelSize == "35M"))
pr_squared_35M <- summary(pmodel_35M)$r.squared
plabel_35M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_35M)))


#ESM2 150M
pmodel_150M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM2" & ModelSize == "150M"))
pr_squared_150M <- summary(pmodel_150M)$r.squared
plabel_150M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_150M)))


#ESM2 650M
pmodel_650M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM2" & ModelSize == "650M"))
pr_squared_650M <- summary(pmodel_650M)$r.squared
plabel_650M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_650M)))

#ESM2 3B
pmodel_3B <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM2" & ModelSize == "3B"))
pr_squared_3B <- summary(pmodel_3B)$r.squared
plabel_3B <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_3B)))


#ESM2 15B
pmodel_15B <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESM2" & ModelSize == "15B"))
pr_squared_15B <- summary(pmodel_15B)$r.squared
plabel_15B <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_15B)))

#ESMC 300M
pmodel_C300M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESMC" & ModelSize == "300M"))
pr_squared_C300M <- summary(pmodel_C300M)$r.squared
plabel_C300M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_C300M)))

#ESMC 600M
pmodel_C600M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "ESMC" & ModelSize == "600M"))
pr_squared_C600M <- summary(pmodel_C600M)$r.squared
#plabel_C600M <- bquote(R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_C600M)))
plabel_C600M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_C600M)))

#Amplify 120M 
pmodel_A120M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "Amplify" & ModelSize == "120M"))
pr_squared_A120M <- summary(pmodel_A120M)$r.squared
plabel_A120M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_A120M)))

#Amplify 350M
pmodel_A350M <- lm(mean_score ~ log(prot_length), data = subset(data_with_size, Model == "Amplify" & ModelSize == "350M"))
pr_squared_A350M <- summary(pmodel_A350M)$r.squared
plabel_A350M <- bquote("Regression " ~ R^2 ~ ": " ~ .(sprintf("%.3f", pr_squared_A350M)))




prot_data_text <- data.frame(
  #labels = c(label_150M, label_650M, label_15B),
  labels = I(list(plabel_ESM1v_650M, plabel_8M, plabel_35M, plabel_150M, plabel_650M, plabel_3B, plabel_15B, plabel_C300M, plabel_C600M, plabel_A120M, plabel_A350M)),
  Model = c("ESM1v", "ESM2", "ESM2", "ESM2", "ESM2","ESM2", "ESM2", "ESMC", "ESMC", "Amplify", "Amplify"),
  ModelSize = c("650M", "8M", "35M", "150M", "650M","3B", "15B", "300M", "600M", "120M", "350M"),
  x = c(130, 460, 460, 460, 460, 460, 460, 180, 180, 180, 180),
  y = c(.72, .75, .75, .75, .75, .75, .75, .83, .83, .76, .76)
)


prot_data_text$ModelSize <- factor(prot_data_text$ModelSize, levels = c("8M", "35M", "120M", "150M", "350M", "650M","3B", "15B", "300M", "600M"))

print(prot_data_text)

```


```{r}

# FOR SUP 9

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, parse(text = paste0("10^", log10(x))), "")
}

data_with_size$ModelSize <- factor(data_with_size$ModelSize, levels = c("8M", "35M", "120M", "150M", "350M", "650M","3B", "15B", "300M", "600M"))

sup9_prot_length <- data_with_size %>%
  filter(Model == 'ESM2') %>%
    ggplot(aes(
      x = prot_length, 
      y = mean_score,
      color = group
      )) +
  scale_x_log10(labels = label_function)+


  geom_smooth(
      aes(group = Model),
      method = 'lm', 
      #formula = y ~ log(x), 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
    geom_point(alpha = 1, size = .9
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .7
  
      )+

    
    labs(x = "Protein Length", 
         y = expression(paste("Cross Validation ", R^2))
         #title = param_vec[1]
         #title = cat("ESM2", param),
         #color = ""
         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        #legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10)
        )+
  
    geom_text(
        data = filter(prot_data_text, Model == "ESM2"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,

        )+
  facet_wrap(~ ModelSize, scales = "free_x")
    


 # ggsave(sup9_prot_length,
 #        filename = "sup9.png",
 #        path = here("figs"),
 #        width = 8.5,
 #        height = 5,
 #        dpi=600)

sup9_prot_length 


```



```{r}
#ESMC 300M

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, trans_format("log10", math_format(10^.x))(x), "")
  }


protlen_ESMC <- data_with_size %>%
  filter(Model == 'ESMC' & ModelSize == '300M') %>%
    ggplot(aes(
      x = prot_length, 
      y = mean_score,
      color = group
      )) +
  
  scale_x_log10(labels = label_function)+


    scale_y_continuous(limits = c(0, 0.85))+

  geom_smooth(
      aes(group = Model),
      method = 'lm', 
      #formula = y ~ log(x), 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
    geom_point(alpha = 1, size = 2
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .8,
      size = .6
  
      )+

    labs(x = "Protein Length", 
         y = NULL,
         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        legend.position = "right",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 20, 0, 0)
        )+
  
    geom_text(
        data = filter(prot_data_text, Model == "ESMC" & ModelSize == "300M"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,

        )


# ggsave(protlen_ESMC,
#        filename = "protlen_ESMC.png",
#        path = here("figs"),
#        width = 8,
#        height = 5,
#        dpi=600)

protlen_ESMC 

```

```{r}

# Amplify prot len

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, parse(text = paste0("10^", log10(x))), "")
}

data_with_size$ModelSize <- factor(data_with_size$ModelSize, levels = c("8M", "35M", "120M", "150M", "350M", "650M","3B", "15B", "300M", "600M"))

protlen_Amplify <- data_with_size %>%
  filter(Model == 'Amplify') %>%
    ggplot(aes(
      x = prot_length, 
      y = mean_score,
      color = group
      )) +
  scale_x_log10(labels = label_function)+


  geom_smooth(
      aes(group = Model),
      method = 'lm', 
      #formula = y ~ log(x), 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
    geom_point(alpha = 1, size = .9
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .7
  
      )+

    
    labs(x = "Protein Length", 
         y = expression(paste("Cross Validation ", R^2))
         #title = param_vec[1]
         #title = cat("ESM2", param),
         #color = ""
         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        #legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10)
        )+
  
    geom_text(
        data = filter(prot_data_text, Model == "Amplify"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,

        )+
  facet_wrap(~ ModelSize, scales = "free_x")
    


 # ggsave(protlen_Amplify,
 #        filename = "protlen_Amplify.png",
 #        path = here("figs"),
 #        width = 8.5,
 #        height = 5,
 #        dpi=600)

protlen_Amplify 


```



# Assemble ESM2, ESMC, Amplify
```{r}
ESM2_assembled <- plot_grid(plot_param,
                  sup9_prot_length,
                  labels = c('A', 'B'), 
                  label_size = 12, 
                  nrow = 2
                  )

# ggsave(ESM2_assembled,
#        filename = "ESM2_assembled.png",
#        path = here("figs"),
#        width = 8,
#        height = 9,
#        dpi=600)

ESM2_assembled

```



```{r}
ESMC_assembled <- plot_grid(plot_param_ESMC,
                  protlen_ESMC,
                  labels = c('A', 'B'), 
                  label_size = 12, 
                  rel_widths = c(1, 1.35)
                  )

# ggsave(ESMC_assembled,
#        filename = "ESMC_assembled.png",
#        path = here("figs"),
#        width = 10,
#        height = 4,
#        dpi=600)

ESMC_assembled

```

```{r}
Amplify_assembled <- plot_grid(plot_param_Amplify,
                  protlen_Amplify,
                  labels = c('A', 'B'), 
                  label_size = 12,
                  nrow = 2
                  #rel_heights = c(1, 1)
                  )

# ggsave(Amplify_assembled,
#        filename = "Amplify_assembled.png",
#        path = here("figs"),
#        width = 9.5,
#        height = 7,
#        dpi=600)

Amplify_assembled

```




# Fig 5
```{r}

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, parse(text = paste0("10^", log10(x))), "")
}

fig_5_datasize <- data_with_size %>%
  filter(Model == 'ESMC' & ModelSize == '600M') %>%
    ggplot(aes(
      x = dataset_size, 
      y = mean_score,
      color = group
      )) +
    

    
  
  geom_smooth(
      #aes(group = Model),
      method = 'lm', 
      #formula = y ~ log10(x), 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
  
  scale_x_log10()+


  
    scale_y_continuous(limits = c(0, 0.85))+
  
    geom_point(alpha = 1, size = 2
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .8,
      size = .6
      )+
    labs(x = "Sample Size", 
         y = expression(paste("Cross Validation ", R^2))
           #expression(paste(R^2)),

         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        legend.position = "none",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10)
        )+
  
    geom_text(
        data = filter(data_text, Model == "ESMC 600M"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,
        
        )

fig_5_datasize

# 
# ggsave(fig_5_datasize,
#        filename = "fig_5_datasize.png",
#        path = here("figs"),
#        width = 8,
#        height = 5,
#        dpi=600)

```


```{r}
#fig 5 second part with protein length

label_function <- function(x) {
    ifelse(log10(x) %% 1 == 0, trans_format("log10", math_format(10^.x))(x), "")
  }


fig5_protlen <- data_with_size %>%
  filter(Model == 'ESMC' & ModelSize == '600M') %>%
    ggplot(aes(
      x = prot_length, 
      y = mean_score,
      color = group
      )) +
  
  scale_x_log10()+


    scale_y_continuous(limits = c(0, 0.85))+

  geom_smooth(
      aes(group = Model),
      method = 'lm', 
      #formula = y ~ log(x), 
      color = "black",      
      se = TRUE, 
      fill = "gray",
      size = 0.3
    ) +
    geom_point(alpha = 1, size = 2
              ) +
    
    geom_errorbar(
      aes(
        ymin = mean_score - std_score, 
        ymax = mean_score + std_score), 
      alpha = .8,
      size = .6
  
      )+

    labs(x = "Protein Length", 
         y = NULL,
         ) +
  guides(color = guide_legend(title = NULL))  +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        legend.position = "right",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 20, 0, 0)
        )+
  
    geom_text(
        data = filter(prot_data_text, Model == "ESMC 600M"),
        mapping = aes(x = x, y = y, label = labels),
        color = "black",
        parse=TRUE,

        )


# ggsave(fig5_protlen,
#        filename = "fig5_protlen.png",
#        path = here("figs"),
#        width = 8,
#        height = 5,
#        dpi=600)

fig5_protlen 

```

```{r}
#fig5 assembly
fig5_assembled <- plot_grid(fig_5_datasize,
                  fig5_protlen,
                  labels = c('A', 'B'), 
                  label_size = 12, 
                  rel_widths = c(1, 1.35)
                  )

# ggsave(fig5_assembled,
#        filename = "fig5_assembled.png",
#        path = here("figs"),
#        width = 10,
#        height = 4,
#        dpi=600)

fig5_assembled

```











# Pisces Data Preprocessing
```{r}
pisces_data <- read.csv(here('results/compression/summary_PISCES_esm2_150M_all_methods.csv'))
pisces_data <- dplyr::select(pisces_data, c('Dataset', 'Compression_method', 'Fold', 'R2_score_test', 'rho_score_test'))


pisces_data

# Convert categorical variables to factors and releveling the mean to be the reference group
pisces_data$Compression_method <- factor(pisces_data$Compression_method)
pisces_data$Compression_method <- relevel(pisces_data$Compression_method, ref = "mean")
pisces_data$Dataset <- factor(pisces_data$Dataset)
pisces_data$Fold <- factor(pisces_data$Fold)

pisces_data

```


```{r}
pisces_model <- lmer(R2_score_test ~ Compression_method + (1 | Dataset) + (1|Fold), data=pisces_data)
summary(pisces_model)
```

```{r}
pisces_glht_results <- summary(glht(pisces_model, linfct=c('Compression_methodbos=0',
                             'Compression_methodiDCT1=0',
                             'Compression_methodiDCT2=0',
                             'Compression_methodiDCT3=0',
                             'Compression_methodiDCT4=0',
                             'Compression_methodiDCT5=0',
                              'Compression_methodmaxPool=0',
                              'Compression_methodpca1=0',
###                              '`Compression_methodpca1-2`=0',
                              'Compression_methodpca2=0',
                              'Compression_methodrbf1=0',
                              'Compression_methodrbf2=0',
                              'Compression_methodsigmoid1=0',
                              'Compression_methodsigmoid2=0')))

pisces_glht_results 

```

```{r}
# Tidy the results
pisces_tidy_glht <- tidy(pisces_glht_results)

pisces_ci <- tidy(confint(pisces_glht_results)) %>% dplyr::select("contrast", "conf.high", "conf.low")

pisces_res_glht = merge(pisces_tidy_glht, pisces_ci, by = 'contrast')



pisces_res_glht$contrast <- gsub("Compression_method", "", pisces_res_glht$contrast)


# Determine significance levels
pisces_res_glht$signif <- ifelse(pisces_res_glht$adj.p.value < 0.001, "***",
                 ifelse(pisces_res_glht$adj.p.value < 0.01, "**",
                 ifelse(pisces_res_glht$adj.p.value < 0.05, "*", "")))



pisces_res_glht
```


# Figure 2B: Pisces lmer
```{r, fig.width=4, fig.height=8, out.width = 100%, out.height = 100%}

pisces_res_glht <- pisces_res_glht %>%
  mutate(contrast = case_when(
    contrast == "bos" ~ "BOS",
    contrast == "maxPool" ~ "Max Pooling",
    contrast == "rbf1" ~ "RBF PC1",
    contrast == "rbf2" ~ "RBF PC2",
    contrast == "sigmoid1" ~ "Sigmoid PC1",
    contrast == "sigmoid2" ~ "Sigmoid PC2",
    contrast == "pca1" ~ "PC1",
    contrast == "pca2" ~ "PC2",
    TRUE ~ contrast
  ))


pisces_plot <- pisces_res_glht %>%
  ggplot(aes(
      y = fct_reorder(contrast, estimate), 
      x = estimate, 
      #color = contrast
      )) +
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, 
                    xmax = conf.high), 
                width = 0.2) +
  theme_minimal() +
  labs(
       y = NULL,
       x = expression(R^2 * " Difference from Mean Pooling"),
       ) +
  geom_vline(xintercept = 0, 
             linetype = "dotted", 
             color = '#000000', 
             linewidth = 1) +

  scale_x_continuous(limit = c(-.85, 0),
                     breaks = seq(-0.8, 0, by = 0.4)
                     ) + 
   theme_minimal_vgrid(12, rel_small = 1) +  
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    plot.background = element_rect(fill = "white", colour = "white"),
    axis.text.y=element_text(size=10),

    panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
    axis.line = element_blank(),
    axis.ticks.x = element_line(color = "black"),
    axis.ticks.y = element_line(color = "black"),
    
    plot.margin = margin(10,10,0,10)
  ) +
  ggtitle("PISCES")

pisces_plot

# ggsave(pisces_plot, 
#        filename = "fig3_iDCT_pisces_v04.png",
#        path = here("figs"),
#        width = 6,
#        height = 6,
#        dpi=600)




  
```







# Figure 3B Pisces multiboxplot

```{r}
multi_mod_pisces <- read.csv(here("results/lassoCV/PISCES/PISCES_all_multi_model.csv"))


multi_mod_pisces <- multi_mod_pisces %>%
  mutate(Model = case_when(
    Model == "esm1v" ~ "ESM-1v",
    Model == "esm2" ~ "ESM-2",
    Model == "esmc" ~ "ESM C",
    Model == "amplify" ~ "AMPLIFY"
  )) %>%
  mutate(
    parameters = case_when(
      parameters == 6.5e8 & Model == "ESM-2" ~ 7.8e8,
      parameters == 6e8 & Model == "ESM C" ~ 5.5e8,
      TRUE ~ parameters
    )
  )


grouped_multi_mod_pisces <- multi_mod_pisces %>%
  filter(Compression_method %in% c('mean')) %>%
  group_by(Dataset, Model, parameters) %>%
  summarise(
    mean_score = mean(get('R2_score_test'), na.rm = TRUE),
    std_score = sd(get('R2_score_test'), na.rm = TRUE),
   
  ) %>%
  mutate(Model = factor(Model, levels = c("ESM-1v", "ESM-2", "ESM C", "AMPLIFY"))) %>%
  ungroup()



data_medians_pisces <- grouped_multi_mod_pisces %>%
  group_by(Model, parameters) %>%
  summarize(
    median_R2 = median(mean_score)
  )



```

```{r}
pisces_all_boxplot <- grouped_multi_mod_pisces %>%
  ggplot(aes(
    x = parameters, 
    y = mean_score, 
    group = interaction(Model, parameters), 
    color = Model,
    fill = Model
    )) +
    geom_line(
      data = data_medians_pisces,
      aes(parameters, median_R2, color = Model),
      inherit.aes = FALSE
    ) +
    geom_boxplot()+
    scale_x_log10(trans = scales::log_trans(),
                     breaks = scales::log_breaks(),
                     labels = trans_format("log10", math_format(10^.x))) +
    scale_color_manual(
      values = colorspace::darken(c(`ESM-1v` = "#000000", `ESM-2` = "#4292C6",
                 `ESM C` = "#228C22", AMPLIFY = "#ff7f0e"), .2)
    ) +
    scale_fill_manual(
      values = c(`ESM-1v` = "#00000080", `ESM-2` = "#9ECAE1",
                 `ESM C` = "#5ad75a", AMPLIFY = "#ffb16c")
    ) +
    labs(x = "", 
         y = expression(paste("Cross Validation ", R^2)), 
         #title = ""
         color = ""
         ) +
    theme_minimal_grid(12, rel_small = 1)+
    theme(
        legend.position = "none",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(10, 10, 0, 10))
  
pisces_all_boxplot
 
# ggsave(pisces_all_boxplot, 
#       filename = "pisces_all_boxplot.png",
#        path = here('figs'),
#        width = 8,
#        height = 4,
#       dpi=600)      



```



# Figure 4 & Sup 4: DMS sampling
```{r}
DMS_sampling <- read.csv(here('results/lassoCV_sampling/lassoCV_sampling_all_models_and_datasets.csv'))

```

```{r}
grouped_DMS_sampling <- DMS_sampling %>%
  mutate(Dataset = case_when(
  Dataset == "BLAT_ECOLX_2015" ~ "BLAT Singles",
  Dataset == "PABP_doubles" ~ "PABP Doubles",
  Dataset == "HIS7" ~ "HIS7 Multiple",
  #TRUE ~ Dataset
  )) %>%
  group_by(Dataset, Sample_size, Model, ModelSize) %>%
  filter(!Sample_size %in% c(32)) %>%
  filter(!Model %in% c("esm1v")) %>%
  summarise(
    mean_R2_score = mean(get('R2_score_test'), na.rm = TRUE),
    std_R2_score = sd(get('R2_score_test'), na.rm = TRUE),
    mean_coeff_score = mean(get('nun_zero_coefs'), na.rm = FALSE),
    std_coeff_score = sd(get('nun_zero_coefs'), na.rm = FALSE)
  ) %>%
  mutate(Model = case_when(
    Model == "esm1v" ~ "ESM-1v",
    Model == "esm2" ~ "ESM-2",
    Model == "esmc" ~ "ESM C",
    Model == "amplify" ~ "AMPLIFY",
    TRUE ~ Model
  )) %>%
  ungroup()


grouped_DMS_sampling$modelname <- paste(grouped_DMS_sampling$Model, grouped_DMS_sampling$ModelSize)

grouped_DMS_sampling <- grouped_DMS_sampling %>%
  mutate(ModelSize = factor(ModelSize, levels = c("8M", "35M", "150M", "650M", "3B", "15B", "300M", "600M", "120M", "350M"))) %>%
  mutate(Dataset = factor(Dataset, levels = c("BLAT Singles", "PABP Doubles", "HIS7 Multiple"))) %>%
mutate(Model = factor(Model, levels = c("ESM-2", "ESM C", "AMPLIFY"))) #%>%
# mutate(modelname = factor(modelname, levels = c("ESM2 8M", "ESM2 35M", "ESM2 150M", "ESM2 650M", "ESM2 3B", "ESM2 15B", "ESMC 300M", "ESMC 600M", "Amplify 120M", "Amplify 350M")))



```


```{r}

#color_dict <- c("150M" = '#9ECAE1','650M' = '#4292C6', "15B" = '#254369')
# color_dict = c("ESM1v 650M" = "#D68F6B", "ESM2 8M" = "#C6DBEF", "ESM2 35M" = "#9ECAE1", "ESM2 150M" = "#6BAED6", "ESM2 650M" = "#4292C6", "ESM2 3B" = "#2171B5", "ESM2 15B" = "#084594", "ESMC 300M" = "#5ad75a", "ESMC 600M" = "#228C22", "Amplify 120M" = "#ffb16c", "Amplify 350M" = "#ff7f0e")

color_dict = c("8M" = "#C6DBEF", "35M" = "#9ECAE1", "150M" = "#6BAED6", "650M" = "#4292C6", "3B" = "#2171B5", "15B" = "#084594", "300M" = "#5ad75a", "600M" = "#228C22", "120M" = "#ffb16c", "350M" = "#ff7f0e")


#5ad75a
DMS_sample_R2_plot <- grouped_DMS_sampling %>%
  ggplot(aes(
      x = Sample_size, 
      y = mean_R2_score,
      color = ModelSize
      )) +
  scale_x_continuous(limits = c(95, 1000000),trans = scales::log_trans(),
                     breaks = scales::log_breaks(),
                     labels = trans_format("log10", math_format(10^.x))
                       ) +
  scale_y_continuous(limits = c(-.3, 1)) +
  geom_errorbar(aes(
      ymin = mean_R2_score - std_R2_score, 
      ymax = mean_R2_score + std_R2_score), 
      width = 0.1,
      alpha = 1,
      #position = position_dodge(width = .2)
      ) +
  geom_line(#position = position_dodge(width = 0.2)
            )+
  geom_point(#position = position_dodge(width = 0.2)
             )+
  scale_color_manual(
    values = color_dict,
    name = "Model Size"
    
    ) +
  
  labs(x= "Sample Size", 
    y = expression(paste("Cross Validation ", R^2)), fill = ""
       ) +
  theme_minimal_grid(12, rel_small = 1)+
  theme(
        legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                   size = .5),
        panel.spacing = unit(.30, "cm"),
      
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(0, 10, 0, 10),
        
        strip.background = element_rect(fill = "gray90", color = "black", size = .5),
  strip.text = element_text(face = "bold", color = "black")
        )+
  facet_grid(Dataset ~ Model, scales = "free")


  # ggsave(DMS_sample_R2_plot,
  #        filename = "DMS_sample_R2_plot.png",
  #        path = here("figs"),
  #        width = 8,
  #       height = 7.5,
  #        dpi=600)


#DMS_sample_R2_plot + facet_grid(cols = vars(Dataset))
DMS_sample_R2_plot


```


```{r}


color_dict = c("8M" = "#C6DBEF", "35M" = "#9ECAE1", "150M" = "#6BAED6", "650M" = "#4292C6", "3B" = "#2171B5", "15B" = "#084594", "300M" = "#5ad75a", "600M" = "#228C22", "120M" = "#ffb16c", "350M" = "#ff7f0e")


sample_coeff_plot <- grouped_DMS_sampling %>%
  ggplot(aes(
      x = Sample_size, 
      y = mean_coeff_score,
      color = ModelSize
      )) +
  scale_x_continuous(limits = c(95, 1000000), trans = scales::log_trans(),
                     breaks = scales::log_breaks(),
                     labels = trans_format("log10", math_format(10^.x))
                       ) +
  scale_y_continuous(limits =c(0, 1350)) +
  #scale_y_continuous(limits = c(NA, 1500)) +
  geom_errorbar(aes(
      ymin = mean_coeff_score - std_coeff_score, 
      ymax = mean_coeff_score + std_coeff_score), 
      width = 0.1,
      alpha = 1,
      #position = position_dodge(width = .2)
      ) +
  geom_line(#position = position_dodge(width = 0.2)
            )+
  geom_point(#position = position_dodge(width = 0.2)
             )+
  scale_color_manual(
    values = color_dict,
    name = 'Model Size'
    ) +
  
  labs(x= "Sample Size", 
    y = "Number of Features"
       ) +
  theme_minimal_grid(12, rel_small = 1)+
  theme(
        legend.position = "top",
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white"),
      

        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                   size = .5),
        panel.spacing = unit(.30, "cm"),
      
        axis.line = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        plot.margin = margin(0, 10, 0, 10),
        
        strip.background = element_rect(fill = "gray90", color = "black", size = .5),
  strip.text = element_text(face = "bold", color = "black")
        )+
  facet_grid(Dataset ~ Model, scales = "free")
        
  
  

# ggsave(sample_coeff_plot,
#        filename = "sample_coeff_plot.png",
#        path = here("figs"),
#        width = 8,
#       height = 7.5,
#        dpi=600)


sample_coeff_plot

```








# Figure 2 Assembly
```{r}
# cowplot for DMS 150M data

double_plot <- plot_grid(final_plot,
                  pisces_plot,
                  labels = c('A', 'B'), 
                  label_size = 12, 
                  rel_widths = c(1.05, 1)
                  )


# ggsave(double_plot,
#       filename = "double_lmer.png",
#       path = here("figs"),
#       width = 8,
#       height = 4,
#       dpi=600)
double_plot

```

# Supplementary Figure 1
```{r}

# mutate df so that it is in the correct form
# use code similar to mega plot above to do all the compression methods
# use cowplot to combine them all

supp_data_mutated = data %>%
  mutate(Compression_method = case_when(
    Compression_method == "bos" ~ "BOS",
    Compression_method == "maxPool" ~ "MaxPooling",
    Compression_method == "rbf1" ~ "RBFPC1",
    Compression_method == "rbf2" ~ "RBFPC2",
    Compression_method == "sigmoid1" ~ "SigmoidPC1",
    Compression_method == "sigmoid2" ~ "SigmoidPC2",
    Compression_method == "pca1" ~ "PC1",
    Compression_method == "pca2" ~ "PC2",
    
    TRUE ~ Compression_method
  )) %>%
  group_by(Dataset, Compression_method) %>%
  summarize(mean_R2_score = mean(R2_score_test, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = Compression_method, values_from = mean_R2_score)

```

```{r}
methods_no_mean <- c("MaxPooling", "iDCT5", 'iDCT4', "PC1", 'RBFPC1', 'SigmoidPC1', "BOS", 'iDCT3', "PC2", 'SigmoidPC2', 'iDCT2', 'RBFPC2', 'iDCT1')
  


scatter_plots_list <- list()

for(method in methods_no_mean){
  dot_color = 'black'

  
  y_label <- ifelse(method == "MaxPooling", "Max Pooling",
                    ifelse(method == "RBFPC1", "RBF PC1",
                           ifelse(method == "SigmoidPC1", "Sigmoid PC1",
                                  ifelse(method == "SigmoidPC2", "Sigmoid PC2",
                                         ifelse(method == "RBFPC2", "RBF PC2", method)))))

  
  
  plot <- supp_data_mutated %>%
    ggplot(aes_string(x = "mean", y = method)) + 
    geom_point(size = 1, color = dot_color) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1)) +
    geom_abline(slope = 1, 
                intercept = 0, 
                linetype = "dashed", 
                color = '#0072B2') +
    labs(x = "Mean Pooling", 
         y = y_label) +
    theme_minimal_grid(12, rel_small = 1) +
    theme(
      panel.background = element_rect(fill = "white", 
                                      colour = "white"),
      plot.background = element_rect(fill = "white", 
                                     colour = "white"),
      legend.position = "none",
      panel.border = element_rect(color = "black", 
                                  fill = NA, 
                                  size = .5
                                  ),
      axis.line = element_blank(),
      axis.ticks.x = element_line(color = "black"),
      axis.ticks.y = element_line(color = "black"),
      plot.margin = margin(8, 8, 8, 8)
    )
  
  scatter_plots_list <- c(scatter_plots_list, list(plot))

}

```


```{r}

final_methods_plot <- plot_grid(plotlist = scatter_plots_list, ncol =3)

# ggsave(final_methods_plot,
#        filename = "sup1_DMS_results_R2.png",
#        path = here("figs"),
#        width = 9,
#        height = 12,
#        dpi=600)

final_methods_plot
```

# Supplementary Figure 2 
```{r}

supp_pisces_mutated = pisces_data %>%
  mutate(Compression_method = case_when(
    Compression_method == "bos" ~ "BOS",
    Compression_method == "maxPool" ~ "MaxPooling",
    Compression_method == "rbf1" ~ "RBFPC1",
    Compression_method == "rbf2" ~ "RBFPC2",
    Compression_method == "sigmoid1" ~ "SigmoidPC1",
    Compression_method == "sigmoid2" ~ "SigmoidPC2",
    Compression_method == "pca1" ~ "PC1",
    Compression_method == "pca2" ~ "PC2",
    TRUE ~ Compression_method
  )) %>%
  group_by(Dataset, Compression_method) %>%
  summarize(mean_R2_score = mean(R2_score_test, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = Compression_method, values_from = mean_R2_score)


methods_no_mean <- c("BOS", "MaxPooling", "iDCT5", "iDCT4", "iDCT3", "PC1", "iDCT2", "SigmoidPC1", "RBFPC1", 'iDCT1', 'PC2', 'SigmoidPC2', 'RBFPC2')
  

pisces_scatter_plots_list <- list()

for(method in methods_no_mean){

  dot_color = 'black'
  y_label <- ifelse(method == "MaxPooling", "Max Pooling",
                    ifelse(method == "RBFPC1", "RBF PC1",
                           ifelse(method == "SigmoidPC1", "Sigmoid PC1",
                                  ifelse(method == "SigmoidPC2", "Sigmoid PC2",
                                         ifelse(method == "RBFPC2", "RBF PC2", method)))))
  
  plot <- supp_pisces_mutated %>%
    ggplot(aes_string(x = "mean", y = method)) + 
    geom_point(size = 1, color = dot_color) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1)) +
    geom_abline(slope = 1, 
                intercept = 0, 
                linetype = "dashed", 
                color = '#0072B2'
                ) +
    labs(x = "Mean Pooling", 
         y = y_label) +
    theme_minimal_grid(12, rel_small = 1) +
    theme(
      panel.background = element_rect(fill = "white", 
                                      colour = "white"),
      plot.background = element_rect(fill = "white", 
                                     colour = "white"),
      legend.position = "none",
      panel.border = element_rect(color = "black", 
                                  fill = NA, 
                                  size = .5),
      axis.line = element_blank(),
      axis.ticks.x = element_line(color = "black"),
      axis.ticks.y = element_line(color = "black"),
      plot.margin = margin(8, 8, 8, 8)
    )
  
  pisces_scatter_plots_list <- c(pisces_scatter_plots_list, list(plot))

}



final_methods_pisces <- plot_grid(plotlist = pisces_scatter_plots_list, ncol =3)

# ggsave(final_methods_pisces,
#        filename = "sup2_pisces_R2.png",
#        path = here("figs"),
#        width = 9,
#        height = 12,
#        dpi=600)

final_methods_pisces



```



# Supp 5

```{r}
#histogram_data <- read.csv(here('results/summary_DMS_all_multi_model.csv'))
histogram_data <- read.csv(here('results/lassoCV/DMS/DMS_all_multi_model_metadata.csv'))

histogram_data <- histogram_data %>%
  group_by(Dataset) %>%
  summarise(
    dataset_size = median(get('dataset_size')),
  ) %>%
  ungroup()


histogram_data$bin <- cut(histogram_data$dataset_size,
                breaks = c(-Inf, 1000, 3200, 10000, 32000, Inf),
                labels = c("<1000", "1000-3200", "3200-10000", "10000-32000", ">32000"),
                right = FALSE) 

histo_plot <- histogram_data %>%
 ggplot(aes(x = bin)) +
  geom_bar(fill = "gray",) +  
  labs(x = "Number of Samples", 
       y = "Number of Datasets", 
       title = "" )+
   theme_minimal_hgrid(12, rel_small = 1) +  
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    plot.background = element_rect(fill = "white", colour = "white"),
    axis.text.x=element_text(size=12),
    axis.text.y=element_text(size=10),

    panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = .5),
    axis.line = element_blank(),
    axis.ticks.x = element_line(color = "black"),
    axis.ticks.y = element_line(color = "black"),
    plot.margin = margin(10,10,0,10)
  ) 
  
histo_plot


# ggsave(histo_plot, 
#        filename = "sup5_DMS_sample_size_v02.png",
#        path = here("figs"),
#        width = 7,
#        height = 5,
#        dpi=600)

```






# Fig 3 multimod assembly
```{r}
fig3_multi_mod <- plot_grid(DMS_all_boxplot, 
                            pisces_all_boxplot, 
                            labels = c('A', 'B'), 
                            ncol = 1, 
                            rel_heights = c(1.1, 1))

ggsave(fig3_multi_mod,
       filename = "fig3_multi_mod.png",
       #path = here("figs"),
       path = 'results/lassoCV/DMS/',
       width = 8,
       height = 8,
       dpi=600)

fig3_multi_mod
```

